Notes on Control Plane Inference

### 2021-03-29 inference.tex
clean up satisfaction

% math
\newcommand{\argmin}{\mathop{\textrm{argmin}}}
\newcommand{\WP}{\textit{wp}}

@@

\paragraph{Boolean Satisfaction}
The satisfaction relation $\pkt,\mu \models_\sigma b$ is defined in
Figure~\ref{fig:bool-satis} in terms of the three kinds of variable contexts,
packets, models, and states. We also use the strict semantics here, most
importantly in the equality rule. Sometimes two expressions are considered equal
if they are both undefined; we do not take this approach, considering the
equality undefined if either argument is undefined.

\begin{figure}
  \[\begin{array}{l>{\iff}cl}
  \pkt,\mu \models_\sigma \FALSE && \textit{pigs fly} \\
  \pkt,\mu \models_\sigma b_1 \Rightarrow b_2 &&
  \pkt,\mu \not\models_\sigma b_1~\textit{or}~
  \pkt,\mu \models_\sigma b_2\\
  \pkt,\mu \models_\sigma e_1 = e_2 &&
  \edenote{e_1}~\pkt~\mu = \edenote{e_2}~\pkt~\mu, \qquad
  \text{(note strict semantics)} \\
  \pkt,\mu \models_\sigma \forall \rho \in t. ~b &&
  \forall r \in \sigma(t.\id).
  \pkt,\sigma,\mu[\rho \mapsto r] \models b
  \end{array}
  \]
  \caption{Satisfaction for $\BExpr$. Write $\pkt \models_\sigma b$ as syntactic
    sugar when the model $\mu$ is empty.}
  \label{fig:bool-satis}
\end{figure}

@@

\begin{figure}[htp]
  \[\begin{array}{l >{\triangleq}cl}
  \multicolumn{3}{l}{\denote{c}^\sigma : \Model \to \Pkt \to \mathcal{P}(\Model \times \Pkt)_\error }\\
    \denote{h.f := e}^\sigma~\mu~\pkt && \{(\mu, \pkt\{h \mapsto \edenote{e}~\pkt~\mu\})\} \\
    \denote{\assume b}^\sigma~\mu~\pkt &&
    \begin{cases}
      \{(\mu,\pkt)\} & \pkt,\mu \models_\sigma b \\
      \{\} & \textit{otherwise} \\
    \end{cases}\\
    \denote{\assert b}^\sigma~\mu~\pkt &&
    \begin{cases}
      \{(\mu,\pkt)\} & \pkt,\mu \models_\sigma b \\
      \error & \textit{otherwise}
    \end{cases} \\
    \denote{\choose \rho t c}^\sigma~\mu~\pkt &&
    \{(\mu'\setminus \rho,\pkt')
    \mid (\mu',\pkt') \in \denote{c}^\sigma\,(\mu\{\rho\mapsto r\})\,\pkt, r \in \sigma(t)\}
    \\
    \denote{c_1;c_2}^\sigma~\mu~\pkt&&
    \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma~\mu'~\pkt' \\
    \denote{c_1 \choice c_2}^\sigma~\mu~\pkt &&
    \denote{c_1}^\sigma~\mu~\pkt \cup \denote{c_2}^\sigma~\mu~\pkt \\
    \denote{t.\apply()}^\sigma~\mu~\pkt &&
    \begin{cases}
      \denote{c[\vec d/\vec x]}^\sigma~\mu~\pkt,
      & \text{where}~\lambda \vec x.~c = r.\action, \vec d = r.\data \\
      & \text{where}~r = \argmin_i\{\pkt,\mu \models_\sigma b_i\} \\
      & \text{where}~b_i = \sigma(t.\id)[i].\matches = t.\keys \\
      \{(\mu,\pkt)\}, & \mathit{otherwise}
    \end{cases}
  \end{array}\]
  \caption{Semantics of Commands}
  \label{fig:com-sem}
  \todo[inline]{$\Model$ then $\Pkt$ everywhere.}
\end{figure}

@@

  \item[$(c = t.\apply())$]
    Let $\sigma$,$\mu$,$\pkt$ be given.
    Let $r_1,\ldots,r_n = \sigma(t)$.
    There are two cases, either the packet hits a row, or hits no row. Explore each case separately.
    \begin{enumerate}
    \item[\textsc{Hit}] In this case, there is some $r_i$ such that $\pkt,\mu
      \models_\sigma r_i.\matches = t.\keys$. By the well-formedness assumption,
      we know that this $r_i$ is unique, and that we can write
      $t.\actions[r_i.\action]$ as $\lambda \vec x. c$, and $|r_i.\data| = |\vec x|$.

@@

      Now, we compute the denotation of $t.\apply()[?]$ to see that it is equivalent to the above.
      The definition of $t.\apply()[?]$ gives us two alternatives. Consider the second, first:
      \[\begin{array}{>{\displaystyle}ll}
        \denote{\assume \forall \rho \in t. \rho.\matches \neq t.\keys;\SKIP}^\sigma~\mu~\pkt \\
        = \denote{\assume \forall \rho \in t. \rho.\matches \neq t.\keys}^\sigma~\mu~\pkt
        & \textrm{\textcolor{red}{Redo with defaults}} \\
        = \begin{cases}
          \{(\mu,\pkt)\}, & \pkt,\mu \models_\sigma \rho.\matches \neq t.\keys \\
          \{\}, &\textit{otherwise}
        \end{cases}
        & \textrm{By definition of }\denote{-}\\
        =  \{\}
        & \textrm{by \textsc{Hit} assumption}
      \end{array}
      \]

@@

    \item[\textsc{Miss}] In this case, there is no $r_i$ such that $\pkt,\mu
      \models_\sigma r_i.\matches = t.\keys$, which means that the LHS of the
      toplevel choice operator should evaluate to $\{\}$ \todo{compute this},
      and the RHS evaluates to $\{\mu,\pkt\}$\todo{compute this}, which is
      precisely the result we get when executing
      $\denote{t.apply()}^\sigma~\mu~\pkt$.

@@

\begin{quote}
  \textbf{Conservative Problem.} Given a command $c \in \Cmd$, compute the weakest test
  $b$ such that if a state $\sigma : \Value \to \Row^*$ the following
  judgement is defined and satisfied \[\emptyset, \emptyset \models_\sigma b,\] then \[\forall \pkt \in \Pkt.~
  \denote{c}^\sigma \pkt \neq \error \]
\end{quote}

@@

\begin{quote}
  \textbf{Flexible Problem.} Given a command $c \in \Cmd$, compute a test
  $b_{\mathit{cp}}$, and a test $b_{\mathit{pkt}}$ such that
  for every packet $\pkt \in \Pkt$ such that
  \[\emptyset, \emptyset \models_\emptyset b_{\mathit{pkt}},\]
  and for every state $\sigma : \Value \to \Row^*$ such that
  \[\pkt, \emptyset \models_\sigma b_{\mathit{cp}},\]
  then \[\denote{c}^\sigma ~\emptyset~\pkt \neq \error \]
\end{quote}

@@

To do so, we need to consider the following dimensions:
\begin{enumerate}[align=left]
\item[\textbf{Keys}.] Should generalize by changing the LHS comparisons to $
  \bigwedge_i[\rho.\keys[i]~\square_{\{=,\neq\}}~?_{v_i}]?$
\item[\textbf{Action Data}.] Should generalize by changing the RHS comparisons to $\bigwedge_i[\rho.\data[i]~\square_{\{=,\neq\}}~?_{v_i}]?$
\item[\textbf{Actions}.] We need to generalize the template across multiple actions. This should be as simple as adding $\rho.\action = i$ on the RHS of each implication above.
\item[\textbf{Sequence}.] I'm not sure how to generalize sequentially composed tables.
  Perhaps a larger catalogue of templates conditioned on structural information:
  if we notice that $t_1$ has a dataflow path to the keys of $t_2$, we should
  generate a join condition template, otherwise we may be able to simply ``and'' the templates.

\item[\textbf{Nondeterminism} ($c_1 \choice c_2$).]. To solve the \textbf{Conservative
  Problem} we simply and the templates together, but to solve the
  \textbf{Flexible Problem} we may be able to constrain the problem (by
  constraining the packets) to either $c_1$ or $c_2$. This would require
  computing a $b_{\textit{pkt}}$.
\end{enumerate}

>>>>

% math
\newcommand{\argmin}{\mathop{\textrm{argmin}}}
\newcommand{\WP}{\textit{wp}}
\newcommand{\satisfy}[3]{({#1,#3}) \models_{#2}}
\newcommand{\notsatisfy}[3]{({#1,#3})\not\models_{#2}}

@@

\paragraph{Boolean Satisfaction}
The satisfaction relation $\satisfy \sigma \pkt \mu b$ is defined in
Figure~\ref{fig:bool-satis} in terms of the three kinds of variable contexts,
packets, models, and states. We also use the strict semantics here, most
importantly in the equality rule. Sometimes two expressions are considered equal
if they are both undefined; we do not take this approach, considering the
equality undefined if either argument is undefined.

\begin{figure}
  \[\begin{array}{l>{\iff}cl}
  \satisfy \sigma \pkt \mu \FALSE && \textit{pigs fly} \\
  \satisfy \sigma \pkt \mu b_1 \Rightarrow b_2 &&
  \notsatisfy \sigma \pkt \mu  b_1~\textit{or}~
  \satisfy \sigma \pkt \mu b_2\\
  \satisfy \sigma \pkt \mu e_1 = e_2 &&
  \edenote{e_1}~\pkt~\mu = \edenote{e_2}~\pkt~\mu, \qquad
  \text{(note strict semantics)} \\
  \satisfy \sigma \pkt \mu \forall \rho \in t. ~b &&
  \forall r \in \sigma(t.\id).
  \satisfy \sigma \pkt {\mu\{\rho \mapsto r\}} b
  \end{array}
  \]
  \todo[inline]{Clarify $\error$ states}
  \todo[inline]{@Nate I know you wanted a standard satisfaction relation, but this was easier to think about in the context of the domain. The difference boils down to a naming convention}
  \caption{Satisfaction for $\BExpr$.}
  \label{fig:bool-satis}
\end{figure}

@@

\begin{figure}[htp]
  \[\begin{array}{l >{\triangleq}cl}
  \multicolumn{3}{l}{\denote{c}^\sigma : \Model \to \Pkt \to \mathcal{P}(\Model \times \Pkt)_\error }\\
    \denote{h.f := e}^\sigma~\mu~\pkt && \{(\mu, \pkt\{h \mapsto \edenote{e}~\pkt~\mu\})\} \\
    \denote{\assume b}^\sigma~\mu~\pkt &&
    \begin{cases}
      \{(\mu,\pkt)\} & \satisfy \sigma \pkt \mu b \\
      \{\} & \textit{otherwise} \\
    \end{cases}\\
    \denote{\assert b}^\sigma~\mu~\pkt &&
    \begin{cases}
      \{(\mu,\pkt)\} & \satisfy \sigma \pkt \mu b \\
      \error & \textit{otherwise}
    \end{cases} \\
    \denote{\choose \rho t c}^\sigma~\mu~\pkt &&
    \{(\mu'\setminus \rho,\pkt')
    \mid (\mu',\pkt') \in \denote{c}^\sigma\,(\mu\{\rho\mapsto r\})\,\pkt, r \in \sigma(t)\}
    \\
    \denote{c_1;c_2}^\sigma~\mu~\pkt&&
    \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma~\mu'~\pkt' \\
    \denote{c_1 \choice c_2}^\sigma~\mu~\pkt &&
    \denote{c_1}^\sigma~\mu~\pkt \cup \denote{c_2}^\sigma~\mu~\pkt \\
    \denote{t.\apply()}^\sigma~\mu~\pkt &&
    \begin{cases}
      \denote{c[\vec d/\vec x]}^\sigma~\mu~\pkt,
      & \text{where}~\lambda \vec x.~c = r.\action, \vec d = r.\data \\
      & \text{where}~r = \argmin_i\{\satisfy \sigma \pkt \mu b_i\} \\
      & \text{where}~b_i = \sigma(t.\id)[i].\matches = t.\keys \\
      \{(\mu,\pkt)\}, & \mathit{otherwise}
    \end{cases}
  \end{array}\]
  \caption{Semantics of Commands}
  \label{fig:com-sem}
  \todo[inline]{$\Model$ then $\Pkt$ everywhere.}
\end{figure}

@@

  \item[$(c = t.\apply())$]
    Let $\sigma$,$\mu$,$\pkt$ be given.
    Let $r_1,\ldots,r_n = \sigma(t)$.
    There are two cases, either the packet hits a row, or hits no row. Explore each case separately.
    \begin{enumerate}
    \item[\textsc{Hit}] In this case, there is some $r_i$ such that $\satisfy \sigma \pkt \mu r_i.\matches = t.\keys$. By the well-formedness assumption,
      we know that this $r_i$ is unique, and that we can write
      $t.\actions[r_i.\action]$ as $\lambda \vec x. c$, and $|r_i.\data| = |\vec x|$.

@@

      Now, we compute the denotation of $t.\apply()[?]$ to see that it is equivalent to the above.
      The definition of $t.\apply()[?]$ gives us two alternatives. Consider the second, first:
      \[\begin{array}{>{\displaystyle}ll}
        \denote{\assume \forall \rho \in t. \rho.\matches \neq t.\keys;\SKIP}^\sigma~\mu~\pkt \\
        = \denote{\assume \forall \rho \in t. \rho.\matches \neq t.\keys}^\sigma~\mu~\pkt
        & \textrm{\textcolor{red}{Redo with defaults}} \\
        = \begin{cases}
          \{(\mu,\pkt)\}, & \satisfy\sigma\pkt\mu \rho.\matches \neq t.\keys \\
          \{\}, &\textit{otherwise}
        \end{cases}
        & \textrm{By definition of }\denote{-}\\
        =  \{\}
        & \textrm{by \textsc{Hit} assumption}
      \end{array}
      \]

@@

    \item[\textsc{Miss}] In this case, there is no $r_i$ such that
      $\satisfy\sigma\pkt\mu r_i.\matches = t.\keys$, which means that the LHS
      of the toplevel choice operator should evaluate to $\{\}$ \todo{compute
        this}, and the RHS evaluates to $\{\mu,\pkt\}$\todo{compute this}, which
      is precisely the result we get when executing
      $\denote{t.apply()}^\sigma~\mu~\pkt$.

@@

\begin{quote}
  \textbf{Conservative Problem.} Given a command $c \in \Cmd$, compute the weakest test
  $b$ such that if a state $\sigma : \Value \to \Row^*$ the following
  judgement is defined and satisfied \[\satisfy\sigma\emptyset\emptyset b,\] then \[\forall \pkt \in \Pkt.~
  \denote{c}^\sigma \pkt \neq \error \]
\end{quote}

@@

\begin{quote}
  \textbf{Flexible Problem.} Given a command $c \in \Cmd$, compute a test
  $b_{\mathit{cp}}$, and a test $b_{\mathit{pkt}}$ such that
  for every packet $\pkt \in \Pkt$ such that
  \[\satisfy \emptyset \pkt \emptyset b_{\mathit{pkt}},\]
  and for every state $\sigma : \Value \to \Row^*$ such that
  \[\satisfy \sigma \pkt \emptyset b_{\mathit{cp}},\]
  then \[\denote{c}^\sigma ~\emptyset~\pkt \neq \error \]
\end{quote}

@@

To do so, we need to extend our templates along the following dimensions:
\begin{enumerate}[align=left]
\item[\textbf{Keys}.] Should generalize by changing the LHS comparisons to $
  \bigwedge_i[\rho.\keys[i]~\square_{\{=,\neq\}}~?_{v_i}]?$
\item[\textbf{Action Data}.] Should generalize by changing the RHS comparisons to $\bigwedge_i[\rho.\data[i]~\square_{\{=,\neq\}}~?_{v_i}]?$
\item[\textbf{Actions}.] We need to generalize the template across multiple actions. This should be as simple as adding $\rho.\action = i$ on the RHS of each implication above.
\item[\textbf{Sequence}.] I'm not sure how to generalize sequentially composed tables.
  Perhaps a larger catalogue of templates conditioned on structural information:
  if we notice that $t_1$ has a dataflow path to the keys of $t_2$, we should
  generate a join condition template, otherwise we may be able to simply ``and'' the templates.

\item[\textbf{Nondeterminism} ($c_1 \choice c_2$).] To solve the \textbf{Conservative
  Problem} we simply and the templates together, but to solve the
  \textbf{Flexible Problem} we may be able to constrain the problem (by
  constraining the packets) to either $c_1$ or $c_2$. This would require
  computing a $b_{\textit{pkt}}$.
\end{enumerate}
